Report on the Second Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE2)

### 2015-06-01 WSSSPE2_report/WSSSPE2_review.tex
editing in intro and conclusions

WSSSPE1 was organized into sessions, each of which was aimed at discussing one
or more of the themes from the call for papers, with a few paper authors invited
to summarize the other papers in the session as a panel, followed by a general discussion
about that theme. The mapping of papers to themes was done by the organizers.

For WSSSPE2, the organizers wanted to increase the interactivity of the
sessions, and to open the process of creating the sessions to the full program
committee, the paper authors, and others who might attend the workshop. In order
to do this, the organizers decided to use a breakout format for two sessions,
and to use an open process to determine the breakout topics. Specifically, Well
Sorted\footnote{\url{http://www.well-sorted.org}} was used in the following
steps: \begin{enumerate} \item Authors were asked to create Well Sorted
``cards'' for the papers. These cards have a title (50 characters maximum) and a
body (255 characters maximum). \item Authors, program committee members, and
members of the WSSSPE mailing list were asked to sort the cards. Each person
drags the cards, one by one, into groups. A group can have as many cards as the
person wants it to have, and it can have whatever meaning that makes sense to
that person. \item Well Sorted produces a set of averages of all the sorts, with
various numbers of card clusters. \end{enumerate}

The organizers then chose a sort that contained five groups and that felt most
meaningful. After that, they decided on themes for the five groups, namely:
\begin{itemize}
\item Exploring Sustainability
\item Software Development Experiences
\item Credit \& Incentives
\item Reproducibility \& Reuse \& Sharing
\item Code Testing \& Code Review.
\end{itemize}

Finally, since some of the papers were not represented by cards in the process,
they were not placed in groups by the peer-grouping system. The authors of
these papers were asked which groups seemed the best for their papers---these
papers were then placed in those groups. Sections~\ref{sec:exploring}-\ref{sec:code_testing}
discuss the breakout groups, including a list of the papers associated with each
group.

@@

\todo{currently just pre-workshop stuff - needs to be updated}

\katznote{add discussion of keynotes, lightning talks, charge to working groups about building plans, etc.}

@@

The WSSSPE2 workshop continues our experiment from WSSSPE1 in how we can
collaboratively build a workshop agenda. The differences in WSSSPE2 from WSSSPE1
are in using an existing service (EasyChair) to handle submissions and reviews,
rather than an ad hoc process, and using an existing service (well-sorted) to
allow collaborative grouping of papers into themes by all authors, reviewers,
and the community, rather than this being done in an ad hoc manner by the
organizers alone.

WSSSPE actively used the online social network Twitter, with hashtag
``\#WSSSPE''. There were substantially more tweets (messages) during the days of
the workshops WSSSPE2, WSSSPE1.1, and WSSSPE1. Out of about 670 tweets as of Apr
18, 2015, more than 225 were about WSSSPE2 and about 180 were posted during the
day of the workshop. Some of the main points and highlights in the meeting are
shown in Table~\ref{tab:tweets}, which summarizes the top \#WSSSPE tweets from
the day of workshop, selected by the metrics that number of retweets or
favourites larger than five and the sum of two measures greater than ten.

After the workshop, one or more additional papers will be created that will
include the discussions at the workshop. These papers will likely have many
authors, and may be submitted to peer-reviewed journals.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgments} \label{sec:acks}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\todo{feel free to add stuff here}

@@

>>>>

The organizers wanted very interactive
sessions, with the process of creating the sessions open to the full program
committee, the paper authors, and others who might attend the workshop. In order
to do this, the organizers used Well
Sorted\footnote{\url{http://www.well-sorted.org}}, with the following
steps:
\begin{enumerate}
\item Authors were asked to create Well Sorted
``cards'' for the papers. These cards have a title (50 characters maximum) and a
body (255 characters maximum).
\item Authors, program committee members, and
members of the WSSSPE mailing list were asked to sort the cards. Each person
drags the cards, one by one, into groups. A group can have as many cards as the
person wants it to have, and it can have whatever meaning that makes sense to
that person.
\item Well Sorted produced a set of averages of all the sorts, with
various numbers of card clusters.
\end{enumerate}

The organizers then chose a sort that contained five groups that felt most
meaningful. After that, they decided on names for the five groups:
\begin{itemize}
\item Exploring Sustainability
\item Software Development Experiences
\item Credit \& Incentives
\item Reproducibility \& Reuse \& Sharing
\item Code Testing \& Code Review.
\end{itemize}

Finally, since some of the papers were not represented by cards in the process,
they were not placed in groups by the peer-grouping system. The authors of
these papers were asked which groups seemed the best for their papers; these
papers were then placed in those groups. Sections~\ref{sec:exploring}-\ref{sec:code_testing}
discuss the breakout groups, including a list of the papers associated with each
group.

@@

The WSSSPE2 workshop continued our experiment from WSSSPE1 in how we can
collaboratively build a workshop agenda, and we began a new experiment in
how to build a series of workshops into an ongoing community activity.

The differences in workshop organization in WSSSPE2 from WSSSPE1
are in using an existing service (EasyChair) to handle submissions and reviews,
rather than an ad hoc process, and using an existing service (well-sorted) to
allow collaborative grouping of papers into themes by all authors, reviewers,
and the community, rather than this being done in an ad hoc manner by the
organizers alone.

WSSSPE actively used the online social network Twitter, with hashtag
``\#WSSSPE''. There were substantially more tweets (messages) during the days of
the workshops WSSSPE2, WSSSPE1.1, and WSSSPE1. Out of about 670 tweets as of Apr
18, 2015, more than 225 were about WSSSPE2 and about 180 were posted during the
day of the workshop. Some of the main points and highlights in the meeting are
shown in Table~\ref{tab:tweets}, which summarizes the top \#WSSSPE tweets from
the day of workshop, selected by the metrics that number of retweets or
favorites larger than five and the sum of two measures greater than ten.

In terms of building community activities, we wanted to focus primarily on
working groups, which we were able to do, as discussed above, but we
also wanted to make sure that attendees felt they had a chance to get their
ideas across to the whole group, which was the purpose of the lightning talks.
Overall, this seemed to be successful at the time, in terms of both the lightning
talks and the breakout groups, and the discussion of sustainability also led
to interesting and useful results, but the challenge that we have discovered
since WSSSPE2 is that it is very hard to continue the breakout groups'
activities.  The WSSSPE2 participants were willing to dedicate their time to
the groups while they were at the meeting, but since then, they have gone
back to their (paid) jobs.  We need to determine how to tie the WSSSPE
breakout activities to people's jobs, so that they feel that continuing them
is a higher priority that it is now, perhaps through funding the participants,
or through funding coordinators for each activity, or perhaps by getting
the workshop participants to agree to a specific schedule of activities during the
workshop.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgments} \label{sec:acks}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\todo{feel free to add stuff here}
